\documentclass[11pt]{article}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage[margin=1in]{geometry}
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{remark}
\newtheorem{remark}{Remark}
\newcommand{\E}{\mathbb{E}}\newcommand{\R}{\mathbb{R}}
\newcommand{\ee}{\mathcal E}\newcommand{\F}{\mathcal F}

\title{Formalization brief: de-opaquing \texttt{lem\_dynkin}\\
(Dynkin martingale proved; bracket by definition, Beth\'e-playbook)}
\date{7 July 2026}

\begin{document}
\maketitle

\noindent\textbf{Goal.} Remove the \texttt{sorry} at \texttt{lem\_dynkin} in
\texttt{TypeDDecouplingEW.lean} (attached). As with the former
\texttt{asepGreenIntegrand}, the conclusion is pinned to OPAQUE objects
(\texttt{dynkinIsMart}, \texttt{dynkinBracket}) --- unprovable by design.
Sanctioned RESTRUCTURING of the \texttt{lem\_dynkin} block only (documented,
fidelity-repair convention; the Bethe/\texttt{asepKernel} precedent): replace
the opaques by real content, prove what is provable, and shrink what remains
to a definitional citation. All other declarations and leaves untouched;
every consumer must still elaborate. New library-clean file
\texttt{TypeDDecouplingDynkin.lean} encouraged. Whole project must build;
standard axioms only.

\section*{Part 1: the martingale property --- PROVED, against Mathlib's
\texttt{Martingale}}

Replace \texttt{dynkinIsMart} by Mathlib's genuine
\texttt{MeasureTheory.Martingale} for an $\R$-indexed filtration
$\F=(\F_t)_{t\ge0}$ on $\Omega$ (a hypothesis of the statement, together with
adaptedness of $t\mapsto\mathrm{proc}\ t$). Faithful hypothesis bundle
(hproc style, documented field-by-field; these are exactly the Markov--Feller
facts the paper's process possesses):
\begin{itemize}
\item[(i)] a semigroup $P:\R_{\ge0}\to(S\to\R)\to(S\to\R)$ with the
\underline{Markov property}: for $0\le s\le t$ and bounded measurable $g$,
$\mu[\,g(\mathrm{proc}\ t)\mid\F_s\,]=(P_{t-s}g)(\mathrm{proc}\ s)$ a.e.;
\item[(ii)] the \underline{Kolmogorov identity}:
$P_tf-f=\int_0^tP_s(Lf)\,ds$ pointwise on $S$ (for the $f$'s in play);
\item[(iii)] boundedness/measurability of $f$, $Lf$, and joint measurability
of $(s,\omega)\mapsto Lf(\mathrm{proc}\ s\ \omega)$, enough for the
conditional Fubini below (state exactly what is used).
\end{itemize}
Then PROVE: $M^f_t:=f(\mathrm{proc}\ t)-f(\mathrm{proc}\ 0)
-\int_0^tLf(\mathrm{proc}\ s)\,ds$ is a martingale:
integrability from (iii); for $s\le t$,
\[
  \mu[M^f_t-M^f_s\mid\F_s]
  =(P_{t-s}f)(\mathrm{proc}\ s)-f(\mathrm{proc}\ s)
  -\int_s^t(P_{u-s}Lf)(\mathrm{proc}\ s)\,du=0
\]
by (i), the conditional Fubini (pull the time integral through the
conditional expectation, justified by (iii)), and (ii) at $t-s$. This is a
tower-property computation --- do it in full.

\section*{Part 2: the bracket --- definitional, with citation}

Mathlib has no continuous-time predictable quadratic covariation, so the
identity $\langle M^f,M^g\rangle_t=\int_0^t\Gamma(f,g)(\mathrm{proc}\ s)\,ds$
cannot even be STATED against a library object. Repair (document prominently):
DEFINE \texttt{dynkinBracket} as that integral (replacing the opaque), so the
second conjunct becomes definitional; the docstring records that identifying
this with the true predictable bracket is the cited classical fact
(Ethier--Kurtz Ch.~4 / Dellacherie--Meyer), exactly as \texttt{asepKernel}'s
identification rests on Sch\"utz. IN ADDITION, prove the honest
$L^2$-level content that IS accessible (this is the part consumers morally
use): the integrated covariance identity
\[
  \E\big[M^f_t\,M^g_t\big]\ =\ \E\Big[\int_0^t\Gamma(f,g)(\mathrm{proc}\ s)\,ds\Big],
  \qquad
  \Gamma(f,g):=L(fg)-fLg-gLf,
\]
from the same bundle (compute both sides via (i)--(ii); polarize from the
$f=g$ case if simpler; add the needed boundedness of $fg$, $L(fg)$ to (iii)).
If the general-$t$ computation is heavy, the identity for $f=g$
(second moment of $M^f$) is the sanctioned minimum.

\section*{Part 3: wiring}

Restate \texttt{lem\_dynkin} with the bundle and the new conclusions
(Martingale $+$ definitional bracket $+$ the proved $L^2$ identity); delete
the opaques \texttt{dynkinIsMart}/\texttt{dynkinBracket}; keep the docstring's
citations and record this as the block's fidelity repair (the opaque-pinned
form was unprovable by design; the new form proves the martingale property
outright and shrinks the citation to the bracket identification). Check and
rewire any consumer. Target count 7 $\to$ 6 (EW file 5 $\to$ 4).

\begin{remark}[hygiene]
(1) Search Mathlib first: \texttt{MeasureTheory.Martingale} ($\R$-indexed),
\texttt{Filtration}, conditional-expectation API
(\texttt{condExp}), Fubini for conditional expectations
(\texttt{MeasureTheory.condExp\_integral}\dots\ find the right form or prove
the needed instance), \texttt{intervalIntegral} measurability lemmas.
(2) The bundle must be SATISFIABLE and faithful: every field is a standard
fact for a Feller jump process with bounded generator; say so per field.
(3) No other EW leaf may be touched; \texttt{thm\_ewmain}'s wiring must
survive verbatim or with the minimal threading.
(4) Report: which parts of Part 2 were proved ($fg$ general or $f=g$ only),
and the final count.
\end{remark}

\end{document}
