\documentclass[11pt]{article}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage[margin=1in]{geometry}
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{remark}
\newtheorem{remark}{Remark}
\newcommand{\E}{\mathbb{E}}\newcommand{\PP}{\mathbb{P}}\newcommand{\R}{\mathbb{R}}
\newcommand{\eps}{\varepsilon}\newcommand{\Sch}{\mathcal{S}}

\title{Formalization brief (Mitoma campaign, 1):\\
the Fr\'echet package for Schwartz space}
\date{10 July 2026}

\begin{document}
\maketitle

\noindent\textbf{Context.} With Aldous's criterion fully discharged, the
project's single remaining \texttt{sorry} is \texttt{thm\_mitoma}
(Mitoma's tightness criterion for $\mathcal S'(\R)$-valued c\`adl\`ag
processes). The campaign will avoid nuclearity and Hermite bases entirely
(both absent from Mathlib) by the Kallianpur--Xiong compact-confinement
formulation; its functional-analytic engine is Banach--Steinhaus on
$\Sch$, and Mathlib's \texttt{WithSeminorms.banach\_steinhaus} and
\texttt{Seminorm.continuous\_iSup} both require
\texttt{[BarrelledSpace $\R$ $\Sch$]} --- an instance Mathlib does not
have, because it lacks \texttt{CompleteSpace} for \texttt{SchwartzMap}
(we verified: no such instance exists anywhere in Mathlib). This task
builds that missing package. It is deliberately \emph{library-grade
Mathlib-general material} (a future upstream candidate, like the Nash and
Kolmogorov--Rogozin files).

\medskip
\noindent\textbf{Goal.} A NEW self-contained file
\texttt{TypeDDecouplingSchwartzFrechet.lean} (Mathlib imports only; no
existing project file is touched; whole project must still build;
standard axioms only; no \texttt{axiom}/\texttt{sorry}). Work over
Mathlib's own \texttt{SchwartzMap E F} ($E,F$ real normed spaces, or
specialize to $\Sch(\R,\R)$ where genuinely necessary --- state each
result at the generality that is natural, and specialize only with a
documented reason). Deliverables, in dependency order:

\begin{itemize}
\item[(1)] \textbf{Countably generated uniformity.}
$\mathcal U(\Sch(E,F))$\texttt{.IsCountablyGenerated} (the uniformity
filter is countably generated) --- from the countable
seminorm family \texttt{schwartzSeminormFamily} ($\mathbb N\times\mathbb
N$-indexed) via \texttt{schwartz\_withSeminorms} and Mathlib's
seminorm-basis description of the uniformity. (Search Mathlib first: a
\texttt{WithSeminorms}-to-countably-generated bridge may already exist
for countable index types.)
\item[(2)] \textbf{Completeness:} \texttt{CompleteSpace $\Sch(E,F)$},
for the EXISTING instances \texttt{SchwartzMap.instUniformSpace}/%
\texttt{instTopologicalSpace} --- do NOT introduce a new metric or
topology and prove completeness for that; the instance must be for
Mathlib's canonical Schwartz uniformity. Route: by (1) and
\texttt{UniformSpace.complete\_of\_cauchySeq\_tendsto} (or its current
name), sequential completeness suffices. For a Cauchy sequence $(f_j)$:
each seminorm $p_{k,n}$ is Cauchy-uniform, so $x^k\,\partial^n f_j$
converges uniformly on $E$; uniform convergence of all iterated
derivatives gives a smooth limit (Mathlib's
\texttt{hasFDerivAt\_of\_tendstoUniformly} family /
\texttt{iteratedFDeriv} limit lemmas); decay bounds pass to the uniform
limit; convergence holds in every seminorm, hence in the Schwartz
topology. This is the analytic heart of the task.
\item[(3)] \textbf{Baire:} \texttt{BaireSpace $\Sch(E,F)$}. Route: (1)
$+$ (2) $+$ Mathlib's metrization of countably generated uniformities
(\texttt{UniformSpace.pseudoMetrizableSpace} or the
\texttt{metrizable\_uniformity} form that preserves the uniformity, so
completeness transfers) $+$ the Baire instance for complete pseudometric
spaces. Take care that the pseudometric induces the SAME uniformity
(otherwise completeness does not transfer); Mathlib's uniformity-level
metrization theorem is the correct tool.
\item[(4)] \textbf{Barrelled:} \texttt{BarrelledSpace $\R$ $\Sch(E,F)$}
--- should be immediate from (3) via
\texttt{BaireSpace.instBarrelledSpace} (the needed
\texttt{IsTopologicalAddGroup}/\texttt{ContinuousConstSMul} instances
exist). If instance-resolution needs T2/separatedness of $\Sch$ anywhere,
note that the seminorm family is point-separating ($p_{0,0}$ is the
sup-norm) and derive the required separation instance.
\item[(5)] \textbf{Sanity corollaries} (these are what campaign task M3
will consume --- they must elaborate, as a check that the instances
actually fire): (a) Banach--Steinhaus for tempered functionals: for any
family $\mathcal F:\iota\to(\Sch(\R,\R)\to L[\R]\ \R)$ with
$\forall\phi,\ \exists C,\ \forall i,\ |\mathcal F_i\phi|\le C$, the
family is uniformly equicontinuous, i.e.\ there is a single continuous
seminorm $q$ on $\Sch$ with $|\mathcal F_i\phi|\le q(\phi)$ for all
$i,\phi$ (via \texttt{WithSeminorms.banach\_steinhaus}; extract the
dominating-seminorm form). (b) A pointwise supremum
$\phi\mapsto\sup_{m}|\mathcal F_m\phi|$ of countably many tempered
functionals, when finite everywhere, is a continuous seminorm (via
\texttt{Seminorm.continuous\_iSup}).
\end{itemize}

\begin{remark}[hygiene]
(1) SEARCH FIRST at every step: the metrization, sequential-completeness,
and seminorm-uniformity bridges likely exist under names we could not
fully confirm; do not rebuild what Mathlib has.
(2) Fallback ladder: full package $\to$ (1)+(2) proved with (3)/(4) as
documented residuals $\to$ honest report. No fabrication; if a step
genuinely resists, report the precise obstruction (e.g.\ a missing
Mathlib metrization lemma) rather than weakening or re-topologizing.
(3) FIDELITY TRAP (repeat of the task-2 lesson): all instances must
attach to Mathlib's canonical \texttt{SchwartzMap} topology/uniformity.
A completeness proof for a homemade metric is worthless to the campaign
and will be rejected in audit.
(4) Report as \texttt{mitoma1\_report.md}: what was proved at what
generality, the exact instance names, which Mathlib lemmas carried
(2) and (3), and any upstream-worthy observations.
\end{remark}

\end{document}
