\documentclass[11pt]{article}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage[margin=1in]{geometry}
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{remark}
\newtheorem{remark}{Remark}
\newcommand{\E}{\mathbb{E}}\newcommand{\PP}{\mathbb{P}}\newcommand{\R}{\mathbb{R}}
\newcommand{\D}{\mathcal D}\newcommand{\eps}{\varepsilon}

\title{Formalization brief (Skorokhod campaign, 7):\\
\texttt{aldous\_tightness} assembled, and \texttt{prop\_aldous} de-opaqued}
\date{9 July 2026}

\begin{document}
\maketitle

\noindent\textbf{Goal.} Two parts. Part A: assemble the packaged
\texttt{aldous\_tightness} from Task 6's proved components (its report
\texttt{skorokhod6\_report.md} is attached and documents the residual
precisely --- read it first). Part B: de-opaque \texttt{prop\_aldous} in
\texttt{TypeDDecouplingEW.lean} (attached) so that it becomes a PROVED
corollary of Part A --- taking the project's real \texttt{sorry} count from
2 to 1. Attached also: \texttt{TypeDDecouplingSkorokhodAldous.lean} (extend
in place) and \texttt{...SkorokhodMeasurable.lean} (the bridge lemmas).
Frozen: Basic/Compact/Complete/Tight/Measurable. Mathlib imports only; whole
project must build; standard axioms only.

\section*{Part A: the assembly (no new analysis --- the report's words)}

Using only proved components: \texttt{consecutive\_crossing\_bound},
\texttt{cadlagModulus\_le\_of\_crossing},
\texttt{crossTime\_osc\_le}, \texttt{isStoppingTime\_crossTime},
\texttt{aldousQ\_mono\_shift}, and the adapter
\texttt{isTightMeasureSet\_of\_bdd\_of\_modulus\_witness}:
\begin{itemize}
\item[(a)] \underline{Crossing-count induction}: from the per-crossing bound,
w.h.p.\ at most $K(\eps,\delta_0)$ crossings occur before time $1$ and no
two are within $\delta$ --- the good event $G$ is a finite conjunction of
\texttt{crossTime}-value conditions, hence measurable (this is why Task 5
built \texttt{measurableSet\_crossTime\_le}).
\item[(b)] \underline{Boundary bookkeeping}: the tail-cell case per Task 6's
report (the truncated Aldous increment at $1$ covers the short terminal
cell; follow the report's stated route exactly).
\item[(c)] \underline{Conclusion}: on $G$, $w'(\delta)\le2\eps$ pathwise
(\texttt{cadlagModulus\_le\_of\_crossing}); feed the witness adapter together
with condition (i) (sup-norm boundedness) to conclude
\texttt{IsTightMeasureSet}. Deliverables: \texttt{aldous\_tightness}
(conditions (i) $+$ (ii) on a sequence of $\D$-valued random elements
$\Rightarrow$ tight laws) and \texttt{aldous\_of\_moment} (via
\texttt{aldousQ\_le\_of\_second\_moment}).
\end{itemize}

\section*{Part B: \texttt{prop\_aldous} de-opaqued (the Bethe/Dynkin
playbook)}

The current \texttt{prop\_aldous} is stated over the OPAQUE predicates
\texttt{aldousTightAt}/\texttt{aldousModulusCond}/\texttt{realTight} with a
schematic $\zeta:\mathbb N\to\R\to\R$ --- unprovable by design and not a
faithful rendering (no probability space, no laws). Restructure (documented
as a fidelity repair, citing the now-real theory):
\begin{itemize}
\item[(d)] RESTATE \texttt{prop\_aldous} as the real theorem: for
$\D$-valued random elements $X_n$ on probability spaces
$(\Omega_n,\PP_n)$ with the sup-norm boundedness condition and the Aldous
stopping-time condition (\texttt{aldousQ} form), the pushforward laws are
tight on $\D$ --- proved by direct application of Part A's
\texttt{aldous\_tightness}. Keep the paper-side docstring (Thm 6.4,
\cite{Aldous}) and note the statement is now the formalized classical
criterion, not a citation.
\item[(e)] The opaques: \texttt{aldousTightAt} and
\texttt{aldousModulusCond} may be deleted IF nothing else consumes them;
\texttt{realTight} is consumed by \texttt{thm\_mitoma} --- LEAVE
\texttt{realTight} and \texttt{thm\_mitoma} exactly as they are (Mitoma
remains the final citation; its de-opaquing is a future decision, not this
task). Verify every consumer of the touched names still elaborates
(\texttt{thm\_ewmain} especially).
\item[(f)] If \texttt{prop\_aldous} turns out to have downstream consumers
that pattern-match its old opaque shape, thread minimally or keep a
deprecated alias with a docstring --- do not weaken \texttt{thm\_ewmain}.
\end{itemize}

\begin{remark}[hygiene]
(1) Fallback: if any (a)--(c) step genuinely resists, deliver what is proved
with the residual documented --- but the report asserts no new analytic
content is needed; treat a failure to assemble as a signal to re-read the
report, not to weaken statements.
(2) Report: the final statement shapes; whether the project's real
\texttt{sorry} count is now 1 (\texttt{thm\_mitoma} only); the consumer
audit for the deleted opaques.
\end{remark}

\end{document}
