\documentclass[11pt]{article}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage[margin=1in]{geometry}
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{remark}
\newtheorem{remark}{Remark}
\newcommand{\E}{\mathbb{E}}\newcommand{\PP}{\mathbb{P}}\newcommand{\R}{\mathbb{R}}
\newcommand{\D}{\mathcal D}\newcommand{\eps}{\varepsilon}\newcommand{\Q}{\mathbb Q}

\title{Formalization brief (Skorokhod campaign, 8):\\
discharging \texttt{prob\_map\_badSet\_le} --- the last Aldous residual}
\date{9 July 2026}

\begin{document}
\maketitle

\noindent\textbf{Goal.} Discharge the single residual
\texttt{prob\_map\_badSet\_le} in
\texttt{TypeDDecouplingSkorokhodAldous.lean} (attached; extend in place),
identified by Task 7's report (\texttt{skorokhod7\_report.md}, attached ---
read it first). Its two missing ingredients, per that report:
\begin{itemize}
\item[(1)] crossing times ITERATED FROM A STOPPING TIME are stopping times
(the frozen \texttt{isStoppingTime\_crossTime} in the Tight module covers
only a deterministic start);
\item[(2)] the terminal-boundary case (an $\eps$-crossing landing within
$\delta$ of time $1$), with the classical level-gap bookkeeping.
\end{itemize}
On success, \texttt{aldous\_tightness}, \texttt{aldous\_of\_moment}, and
\texttt{prop\_aldous} lose their transitive \texttt{sorryAx} and the
project's real \texttt{sorry} count becomes 1 (\texttt{thm\_mitoma} only).
Frozen: Basic/Compact/Complete/Tight/Measurable (Tight and EW attached for
reference; new lemmas live in the Aldous file). Mathlib imports only; whole
project must build; standard axioms only.

\section*{Item (1): crossing after a stopping time}

For the coordinate process on $\D$ with its right-continuous coordinate
filtration, and a stopping time $\tau$: the first $\eps$-crossing AFTER
$\tau$,
$\sigma:=\inf\{t>\tau:|X_t-X_\tau|\ge\eps\}$ (clamped as in
\texttt{crossSeq}), is again a stopping time. Route: the same
rational-reduction as the fixed-start case ---
$\{\sigma\le t\}=\bigcup_{q\in\Q\cap[0,t]}\big(\{\tau\le q\}\cap
\{\exists\,r\in\Q\cap(q,t]:|X_r-X_{\text{(value at }\tau)}|\ge\eps'\}\big)$-style
--- with the value $X_\tau$ handled through the stopping $\sigma$-algebra
(Mathlib's \texttt{IsStoppingTime.measurableSpace} API; \texttt{stoppedValue}
measurability via the joint measurability \texttt{measurable\_eval\_shift}
of Task 6). Mind right-continuity for the $\ge$ vs $>$ threshold, exactly as
in \texttt{crossTime\_le\_iff}. Conclude by induction: every
\texttt{crossSeq} iterate is a stopping time, so Task 6's
\texttt{consecutive\_crossing\_bound} applies at each step (its hypothesis
is a stopping time, which is what Item (1) supplies).

\section*{Item (2): the terminal boundary}

The event ``some crossing lands in $(1-\delta,1]$'' per Task 7's report:
handle by the truncated Aldous increment at $1$ (the \texttt{aldousQ}
truncation \texttt{min}$(\tau+\delta,1)$ was designed for this; Task 5's
convention counterexample dictates the tail-long treatment). SANCTIONED:
the constants and threshold levels in \texttt{prob\_map\_badSet\_le} (and,
if forced, the numeric constants inside \texttt{aldous\_tightness}'s proof
--- NOT its statement's meaning) may be adjusted to whatever the honest
bookkeeping yields; Task 7's report explicitly anticipated this. Any change
must keep \texttt{aldous\_tightness}/\texttt{prop\_aldous} conclusions
semantically identical (tightness from boundedness $+$ vanishing Aldous
quantity); document every adjustment.

\begin{remark}[hygiene]
(1) Fallback ladder: full discharge $\to$ Item (1) proved with Item (2) as a
sharpened documented residual $\to$ honest report. No fabrication; treat a
statement that cannot be proved as a statement to fix (documented), not to
force.
(2) Report: the final form of \texttt{prob\_map\_badSet\_le}; whether
constants changed; confirmation that \texttt{prop\_aldous} and
\texttt{aldous\_tightness} are now \texttt{sorryAx}-free; the real count
(target: 1).
\end{remark}

\end{document}
