\documentclass[11pt]{article}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage[margin=1in]{geometry}
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{remark}
\newtheorem{remark}{Remark}
\newcommand{\E}{\mathbb{E}}\newcommand{\PP}{\mathbb{P}}\newcommand{\R}{\mathbb{R}}
\newcommand{\D}{\mathcal D}\newcommand{\eps}{\varepsilon}\newcommand{\Leb}{\mathrm{Leb}}

\title{Formalization brief (Skorokhod campaign, 9):\\
the genuine Aldous architecture --- replacing the refuted \texttt{badSet} route}
\date{10 July 2026}

\begin{document}
\maketitle

\noindent\textbf{Context.} Task 8's counterexample (see
\texttt{skorokhod8\_report.md}, attached --- read it first) refuted
\texttt{prob\_map\_badSet\_le}: no $\delta$-independent constant exists,
because the ``first bad crossing index'' is future-dependent and the honest
union bound costs a factor $\lceil 1/\delta\rceil$. That analysis was
correct. We have since obtained the \emph{original proof} (D.~Aldous,
``Stopping times and tightness,'' \emph{Ann.\ Probab.}~6 (1978) 335--340,
pp.~338--340) and transcribe its actual mechanism below. The union-bound
cost is paid by a \textbf{two-scale invocation} of the Aldous hypothesis:
once at scale $\delta$ with tolerance $\eta_1$ (for the crossing count), and
once at a finer scale $\sigma\le\delta$ with tolerance $\eta_2/q$ where
$q\approx 2/\delta$ is the number of crossing indices (for the gap
separation). No index selection is ever adapted-filtered: each crossing time
is itself a stopping time and receives its own direct bound.

\medskip
\noindent\textbf{Goal.} In \texttt{TypeDDecouplingSkorokhodAldous.lean}
(attached; edit in place):
\begin{itemize}
\item DELETE the false \texttt{prob\_map\_badSet\_le} (and
\texttt{badSet}/\texttt{measurableSet\_badSet} if nothing else consumes
them --- audit consumers) and REPROVE \texttt{aldous\_tightness} by the
architecture below. The \emph{statements} of \texttt{aldous\_tightness},
\texttt{aldous\_of\_moment}, and \texttt{prop\_aldous} (in the EW file) are
UNCHANGED --- we have verified that the existing hypothesis
\texttt{hald} ($\forall\eps>0\ \forall\eta>0\ \exists\delta>0\ \forall i,\
\texttt{aldousQ}\ \delta\ \eps\le\eta$) already supplies both scales, and
that \texttt{aldousQ}'s supremum over deterministic shifts $\delta'\le d$
is exactly Aldous's condition (13). On success the project's real
\texttt{sorry} count is 1 (\texttt{thm\_mitoma} only).
\end{itemize}
Frozen: Basic/Compact/Complete/Tight/Measurable (attached for reference;
new lemmas live in the Aldous file). Mathlib imports only; whole project
must build; standard axioms only.

\section*{The architecture (Aldous 1978, (13)--(25), transcribed)}

Fix $\eps>0$, target failure probability $\eta$. Write $X_t$ for the
coordinate evaluation and recall \texttt{flatR}: $X_{(\tau+s)\wedge 1}=
X_{\tau+s}$ in effect, so the \texttt{aldousQ} truncation matches Aldous's
$[0,2]$-extension for free.

\paragraph{Step 1 (have already).} The crossing sequence $T_0=0$,
$T_{i+1}=$ first $t>T_i$ with $|X_t-X_{T_i}|\ge 2\eps$ (clamped as in
\texttt{crossSeq}); every $T_i$ is a stopping time ---
\texttt{isStoppingTime\_crossSeq} from Task 8 stands and is the backbone.

\paragraph{Step 2 (shift-average bound; Aldous (23)--(24)).} For a stopping
time $\tau\le1$ and $\delta>0$ let
\[
A_\delta(\tau):=\Big\{\omega:\ \Leb\{s\in[0,2\delta]:
|X_{(\tau+s)\wedge1}-X_\tau|\ge\eps\}\ \ge\ \tfrac{\delta}{2}\Big\}.
\]
Then $\PP(A_\delta(\tau))\le 4\,\texttt{aldousQ}(2\delta,\eps)$: the fiber
measure $L(\omega)$ has $\E L=\int_0^{2\delta}\PP(|X_{(\tau+s)\wedge1}
-X_\tau|\ge\eps)\,ds\le 2\delta\cdot\texttt{aldousQ}(2\delta,\eps)$ by
\texttt{le\_aldousQ\_of\_stoppingTime} (each deterministic $s\le2\delta$ is
admissible in the supremum), and Markov at level $\delta/2$ finishes. Tools:
Fubini via \texttt{MeasureTheory.intervalIntegral\_integral\_swap} (or
\texttt{Measure.prod} + \texttt{lintegral\_lintegral\_swap}); joint
measurability of $(\omega,s)\mapsto X_{(\tau(\omega)+s)\wedge1}(\omega)$
from Task 6's \texttt{measurable\_eval\_shift} plus stopped-value
measurability as in Task 8. \emph{No abstract conditional expectation is
needed anywhere} --- Aldous's ``$\PP(\cdot\mid\mathcal F)\ge\frac14$''
device is literally this explicit fiber average.

\paragraph{Step 3 (pigeonhole; Aldous (21)--(22)).} Deterministic lemma: if
$t_2-t_1<\delta$ and for $j=1,2$ the bad sets
$\{s\in[0,2\delta]:|f(t_j+s)-f(t_j)|\ge\eps\}$ each have Lebesgue measure
$<\delta/2$, then --- since the windows $[t_1,t_1+2\delta]$ and
$[t_2,t_2+2\delta]$ overlap in an interval of length
$t_1+2\delta-t_2>\delta$, and the two bad sets remove $<\delta$ of it ---
there is a common good time $\theta_0$, whence
$|f(t_2)-f(t_1)|<2\eps$. Contrapositive, applied at $t_1=T_{i-1}$,
$t_2=T_i$ (where $|X_{T_i}-X_{T_{i-1}}|\ge2\eps$ by construction):
\[
\{T_i<1,\ T_i<T_{i-1}+\delta\}\ \subseteq\
A_\delta(T_{i-1})\cup A_\delta(T_i),
\]
so by Step 2, $\PP(T_i<1,\ T_i<T_{i-1}+\delta)\le
8\,\texttt{aldousQ}(2\delta,\eps)$ \emph{for each fixed} $i$. (Handle the
window clipping at time $1$ via \texttt{flatR}; on $\{T_i<1\}$ both
crossing values are genuine.)

\paragraph{Step 4 (two scales and assembly; Aldous (14)--(17), (25)).}
Instantiate \texttt{hald} twice:
\begin{itemize}
\item at $(\eps',\eta_1)$ obtain $\delta$ with
$\texttt{aldousQ}(2\delta,\eps')\le\eta_1$ uniformly in $i$; fix an integer
$q$ with $q\delta>2$ (e.g.\ $q=\lceil2/\delta\rceil+1$);
\item at $(\eps',\eta_2/q)$ obtain $\sigma$, and replace $\sigma$ by
$\min(\sigma,\delta)$ using \texttt{aldousQ\_mono\_shift}.
\end{itemize}
Then:
\begin{itemize}
\item[(a)] \underline{Separation (Aldous (16),(25))}: union over
$i=1,\dots,q$ of Step 3 at scale $\sigma$ gives
$\PP(\exists i\le q:\ T_i<1,\ T_i<T_{i-1}+\sigma)\le 8\eta_2$. The factor
$q$ is absorbed by the $\eta_2/q$ tolerance --- this is precisely the
device whose absence made Task 8's statement false.
\item[(b)] \underline{Crossing count (Aldous (15))}: with $Y_i:=T_i\wedge1
-T_{i-1}\wedge1$ and $B:=\{T_q<1\}$, event-conditioning only
($\E[Y\,\mathbf1_B]$ manipulations, no filtration conditioning):
$\E[Y_i\mathbf1_B]\ge\delta\big(\PP(B)-\PP(T_i<1,\,T_i<T_{i-1}+\delta)\big)
\ge\delta(\PP(B)-8\eta_1)$ by Step 3 at scale $\delta$; summing,
$\PP(B)\ge\E[(T_q\wedge1)\mathbf1_B]\ge q\delta(\PP(B)-8\eta_1)$, and
$q\delta>2$ forces $\PP(T_q<1)\le 16\eta_1$.
\item[(c)] \underline{Conclusion (Aldous (17))}: on the good event
$\{T_q\ge1\}\cap\{\text{all gaps}\ge\sigma\}$, the path has fewer than $q$
$2\eps'$-crossings before $1$, no two within $\sigma$, so
\texttt{cadlagModulus\_le\_of\_crossing} gives $w'(X,\sigma)\le c\,\eps'$
pathwise; failure probability $\le16\eta_1+8\eta_2$. Choose
$\eps',\eta_1,\eta_2$ from the target $(\eps,\eta)$ of the witness adapter
\texttt{isTightMeasureSet\_of\_bdd\_of\_modulus\_witness} and conclude
\texttt{aldous\_tightness}; \texttt{aldous\_of\_moment} and
\texttt{prop\_aldous} then clear their transitive \texttt{sorryAx}
automatically (verify with an axiom audit).
\end{itemize}
SANCTIONED: all numeric constants ($16$, $8$, $\tfrac{\delta}{2}$, the
$c$ in $w'\le c\,\eps'$, the $\eps'/\eta_j$ bookkeeping) are free to be
whatever the honest accounting yields; the theorem \emph{statements} must
not change meaning.

\begin{remark}[hygiene]
(1) Fallback ladder: full discharge $\to$ Steps 2--3 proved with Step 4
assembly as a sharpened, documented residual (stated in the CORRECT
two-scale form, never the refuted single-scale form) $\to$ honest report.
No fabrication; a statement that cannot be proved is a statement to fix
(documented), not to force.
(2) Docstrings: record that the Task-8 route was refuted by counterexample
and this file implements Aldous's original two-scale argument.
(3) Report: final lemma names and constants; consumer audit for the deleted
\texttt{badSet} items; confirmation that \texttt{aldous\_tightness},
\texttt{aldous\_of\_moment}, \texttt{prop\_aldous} are \texttt{sorryAx}-free;
the real count (target: 1, \texttt{thm\_mitoma} only).
\end{remark}

\end{document}
